% max 10 pages

 \subsection{Ethical and/or legal aspects of the project}
 \label{sec:ethical}

\subsubsection{General ethical aspects}

Ethical and legal considerations are not applicable to the proposed project since we 
will not conduct experiments on humans, animals, or genetic resources, nor will we  produce results that can be intentionally misused to cause harm. 

We will conduct user studies in WP6 of the project to evaluate the the effectiveness and efficiency of the discovered process representation. Such user studies are highly common in the context of computer science research. Participation in these studies will be fully voluntarily, participants will have the possibility to opt out at any moment, and we will not collect any personal data as defined by the Ethics Committee of the University of Mannheim\footnote{\url{https://www.uni-mannheim.de/en/about/organization/bodies-and-committees/committees-and-councils/ethics-committee}}. Given that we will not collect any personal data, an ethics vote is not required according to the committee's guidelines. 

\subsubsection{Descriptions of proposed investigations involving humans, human materials or identifiable data}

Not applicable.

\subsubsection{Descriptions of proposed investigations involving experiments on animals}

Not applicable.

\subsubsection{Descriptions of projects involving genetic resources (or associated traditional knowledge) from a foreign country}

Not applicable.

\subsubsection{Explanations regarding any possible safety-related aspects (“Dual Use Research of Concern; foreign trade law)}

Not applicable.

\subsection{Employment status information}

Prof. Dr. Henrik Leopold, Associate Professor at the Kühne Logistics University, permanent position.\\[1em]
Prof. Dr. Han van der Aa, Junior Professor (W1) at the University of Mannheim with a contract until March 2026.

\subsection{First-time proposal data}

Not applicable.

\subsection{Composition of the project group}

The project group is composed of Prof.\ Henrik Leopold and Prof.\ Han van der Aa:


\textbf{Prof. Dr. Henrik Leopold} - Henrik Leopold is a tenured Associate Professor at the K\"uhne Logistics University (KLU) and senior researcher at the Hasso Plattner Institute (HPI) at the Digital Engineering Faculty, University of Potsdam. Before joining KLU/HPI in February 2019, he held positions as an Assistant Professor at the Vrije Universiteit Amsterdam (February 2015 – January 2019) and WU Vienna (April 2014 – January 2015) as well as a postdoctoral research fellow at the Humboldt University of Berlin (July 2013 – March 2014). In July 2013, he obtained a PhD degree (Dr. rer. pol.) in Information Systems from the Humboldt University of Berlin. For his thesis he received the TARGION Dissertation Award 2014 for the best doctoral thesis in the field of Information Management and the runner-up of the McKinsey Business Technology Award 2013. Henrik Leopold's research is concerned with leveraging technology from the field of artificial intelligence to develop automated techniques for process analysis and process mining. The results of his research have been published in over 100 publications in books, book chapters, journals, conferences, workshops, and reports. Among others, his research has been published in the journals IEEE Transactions on Knowledge and Data Engineering, IEEE Transactions on Software Engineering, ACM Transactions on Management Information Systems, Decision Support Systems, and Information Systems. 

\textbf{Prof. Dr. Han van der Aa} - Han van der Aa is a Junior Professor (W1) in the School of Business Informatics at the University of Mannheim, where he heads the research group on process analytics since April 2020. Before that, he was an Alexander von Humboldt Fellow, working as a postdoctoral researcher in the Department of Computer Science at the Humboldt-Universität zu Berlin (May 2018 - March 2020). He obtained a PhD in computer science from the Vrije Universiteit Amsterdam in January 2018. His research interests include business process modeling, process mining, natural language processing, and complex event processing. 
The results of his research have so far resulted in close 50 publications, including articles in 
renowned international journals such as IEEE TKDE, DSS, and Information Systems, as well as at the CAISE, BPM, SIGMOD, and COLING conferences.
He is a PC member of established conferences such as BPM, ICPM, and DEBS.


The applicants have a history of successful collaboration. While they share a similar background with respect to the application of natural language processing in process analysis and mining, they  have complementary skills and knowledge, as described in the context of their preliminary work in \autoref{sec:preliminarywork}.

 \subsection{Researchers in Germany with whom you have agreed to cooperate on this project}
\label{sec:collab:germany}

We agreed to cooperate on this project with three researchers from Germany: Prof. Dr. Stefanie Rinderle-Ma, Prof. Dr. Matthias Weidlich, and Prof. Dr. Simone Ponzetto. 

\textbf{Prof. Dr. Stefanie Rinderle-Ma}: Stefanie Rinderle-Ma is a full professor for Information Systems and Business Process Management at the Department of Informatics, Technical University of Munich, Germany. Before Stefanie worked as full professor at the University of Vienna, Austria, where she led the Research Group on Workflow Systems and Technology. Stefanie‘s research interests include flexible and distributed process technology, production intelligence, process mining, and digitalized compliance management.
 
\textbf{Prof. Dr. Matthias Weidlich}: Matthias Weidlich is a full professor at the Department of Computer Science at Humboldt- Universi\"at zu Berlin and has been an Emmy Noether Research group leader at the same institute. Before that, he held positions at the Department of Computing at Imperial College London and at the Technion - Israel Institute of Technology. He holds a PhD from the Hasso Plattner Institute (HPI), University of Potsdam. His research focuses on process-oriented and event-driven systems and his results appear regularly in the premier conferences (VLDB, SIGMOD) and journals (TKDE, Inf. Sys., VLDBJ) in the field.

\textbf{Prof. Dr. Simone Ponzetto}: Simone Ponzetto is a professor of Information Systems at the University of Mannheim, where he leads the Natural Language Processing (NLP) and Information Retrieval group. His main research interests lie in the areas of knowledge acquisition, text understanding, and the application of NLP methods for research in the (digital) humanities and (computational) social sciences. Simone regularly serves as area chair and program committee member of *ACL and (IJC/AA)AI conferences and is an editorial board member of the Artificial Intelligence Journal and Journal of Natural Language Engineering. He is (co-)author and (co-)editor of over 100 refereed papers in scientific journals, books, and conference proceedings.

We have completed and/or ongoing research collaborations with all three researchers.
With Prof. Dr. Stefanie Rinderle-Ma, we worked, among others, on extracting process information from textual resources in the context of compliance checking \cite{winter2020assessing}. 
With Prof. Dr. Matthias Weidlich, we addressed various topics related to process analysis \cite{leopold2012probabilistic,weidlich2013predicting}, process mining \cite{van2020partial,fahrenkrog2020pripel}, and low level event data \cite{zhao2021eires,busany2020interval}.
With Prof. Dr. Simone Ponzetto, we have on ongoing cooperation on  the extraction of process information from text. 
%
Based on this experience, we are confident that all three can provide valuable input for this project. With respect to the work packages, we aim to primarily collaborate with Prof. Rinderle-Ma on WP1 and WP6, with Prof. Weidlich on WP2 and WP3, and with Prof. Ponzetto on WP1 and WP5.


\subsection{Researchers abroad with whom you have agreed to cooperate on this project}
\label{sec:collab:abroad}

Outside Germany, we agreed to cooperate with Prof. Dr. Josep Carmona and Dr. Chiara Ghidini.

\textbf{Prof. Dr. Josep Carmona}: Josep Carmona is an Associate Professor at Universitat Politècnica de Catalunya (UPC). He received a PhD at the same university in 2004, under the supervision of Prof. Jordi Cortadella. His research interests include formal methods and concurrent systems, data and process science, business intelligence and business process management, and natural language processing. He has co-authored numerous research papers and organized various conferences and workshops. He is a founder of the International Conference on Process Mining, where he is part of the Steering Committee, and a member of the IEEE Task Force on Process Mining.

\textbf{Dr. Chiara Ghidini}: Chiara Ghidini is a senior research scientist at Fondazione Bruno Kessler within the Data and Knowledge Management (DKM) research unit. Her work in the area of distributed knowledge representation is well known and internationally recognized and she has published more than 100 conference and journal papers on the topics. She has served on many organising and program committees for conferences and workshops in the areas of multi-agent systems and the semantic web. She has been involved in a number of international research projects, including the EU-funded projects APOSDLE and Organic Lingua and is currently leading the interdisciplinary SHELL project on procedural and ontological knowledge acquisition and evolution at the Fondazione Bruno Kessler. 

We have already worked with Prof. Carmona in the past, primarily on the intersection between NLP and process analysis \cite{sanchez2018aligning,van2018challenges}.  With Dr. Ghidini, we have an ongoing collaboration on the use of deep learning for the analysis of textual process descriptions. 
Against this background, we believe that cooperations with Prof. Carmona and Dr. Ghidini will be a valuable addition to this project, primarily with respect to work packages WP1, WP2, and WP5. 

\subsection{Researchers with whom you have collaborated scientifically within the past three years}


The researchers listed above are also ones we have collaborated with over the last three years. We plan to use this project to continue and deepen these collaborations. Other research collaborations in the context of process mining, process analysis, and natural language processing have been conducted with the researchers below.

\textbf{Prof. Dr. Henrik Leopold}

\begin{itemize}
\item Prof. Dr. Hajo Reijers (Utrecht University, NL): Process mining and analysis
\item Prof. Dr. Mathias Weske (Hasso Plattner Institute, DE): Process mining and automation
\item Prof. Dr. Flávia Maria Santoro (Universidade do Estado do Rio de Janeiro, BR): Process modeling and analysis
\end{itemize}

\textbf{Prof. Dr. Han van der Aa}

\begin{itemize}
\item Prof Dr. Avigdor Gal (Israel Institute of Technology, IL): Process mining
\item Prof. Dr. Jan Mendling (Humboldt-Universit\"at zu Berlin, DE): NLP-based process analysis 
\item Prof. Dr. Heiner Stuckenschmidt (University of Mannheim, DE): AI for process analysis
\item Dr. Fabrizio Maggi (Free University of Bozen-Bolzano, IT): Declarative process analysis
\item Dr. Adela del-R\'{i}o-Ortega (Universidad de Sevilla, ES): Process performance analysis
\end{itemize}


\subsection{Project-relevant cooperation with commercial enterprises}

We do not plan to cooperate with a commercial enterprise in the context of this project. 
%However, as discussed earlier, we plan to reuse an event log we are currently creating together with the German process mining software vendor Lana Labs\footnote{https://lanalabs.com/} in the research project \textit{AutoDrivE} (Automatic Derivation of Event Logs)\footnote{https://www.nwo.nl/onderzoek-en-resultaten/programmas/open+technologieprogramma/projecten/2018/2018-16672/} funded by the Dutch Research Council (NWO).

%As discussed earlier, we plan to cooperate with the German process mining software vendor Lana Labs\footnote{https://lanalabs.com/}. The main goal of the cooperation is to jointly create an additional data set for the evaluation of the proposed technique. We have already collected experience with Lana Labs as a project partner. In the research project \textit{AutoDrivE} (Automatic Derivation of Event Logs) funded by the Dutch Research Council (NWO), we are cooperating on the topic of automated event log extraction from transactional databases. We are, therefore, confident that also the collaboration in the context of this project will be successful.   

\subsection{Project-relevant participation in commercial enterprises}

We do not have any project-relevant participation in commercial enterprises.  

\subsection{Scientific equipment}

We will not need any further scientific equipment. 

\subsection{Other submissions}

We currently have no other proposal submissions under review for this project.  

\subsection{Other information}
