 \subsection{Basic Module}
\subsubsection{Funding for Staff}
\label{sec:staff}

We apply for two doctoral researchers (TV-L 13) for the time of 2 years to conduct the work planned for this project (24 PM). Each position will be assigned to a PhD student who will be responsible for one work stream.

\begin{funds}[funding for staff]
\positionmul{Doctoral researcher (WS1 - Van der Aa), TV-L 13, 24 months}{6175}{24}
\positionmul{Doctoral researcher (WS2 - Leopold), TV-L 13, 24 months}{6175}{24}
\end{funds}

\subsubsection{Direct Project Costs}

\subsubsubsection{Equipment up to \euro10,000, Software and Consumables}

We will not require additional hardware or software. The required infrastructure will be provided by the K\"uhne Logistics University and the University of Mannheim.

\subsubsubsection{Travel Expenses}
\label{sec:costs:travel}
The results of the project will be published and presented at national and international conferences. We plan to visit two conferences per year. Assuming an average of \euro1500 per conference, we will need 2 x 2 \euro1500 = \euro6000 per PhD student, i.e. \euro12000 in total. We plan to conduct two joint project meetings per year: one in Hamburg and one in Mannheim. This can be covered by \euro500 per trip, which means we need \euro2000 for project meetings. We further plan to visit two of our national and one of our international project partners over the course of the project. The trips to the national partners can be covered by \euro500 and the international partners by \euro1000. This sums up to \euro4000 for the entire project. A summary and the overall total required for traveling expenses are listed below. \\

\begin{funds}[travel expenses]
\positionmul{Conference visits (2 years x 2 visits x 2 researchers)}{1500}{8}
\positionmul{Project meetings (2 years x 2 visits)}{500}{4}
\positionmul{Visit national project partner (2 partners x 1 visit x 2 researchers)}{500}{4}
\positionmul{Visit international project partners (1 partner x 1 visit x 2 researchers)}{1000}{2}
\end{funds}

\subsubsubsection{Visiting Researchers}

As explained in Sections \ref{sec:collab:germany} and \ref{sec:collab:abroad}, we plan to cooperate with several national as well as international researchers. Costs of mutual visits are listed in above in Section \ref{sec:costs:travel}.

\subsubsubsection{Expenses for Laboratory Animals}

Not applicable. 

\subsubsubsection{Other Costs}

There will not be any other costs than the positions listed above. 

\subsubsubsection{Project-related publication expenses}

We do not expect any publication expenses besides the conference fees, which we already included in the travel expenses in Section \ref{sec:costs:travel}. The total for direct projects costs is listed below. 

\subsubsection{Instrumentation}

Not applicable. 

\subsection{Module Temporary Position for Principal Investigator}

Not applicable. 

\subsection{Module Replacement Funding}

Not applicable. 

\subsection{Module Temporary Clinician Substitute}

Not applicable. 

\subsection{Module Mercator Fellows}

Not applicable. 

\subsection{Module Workshop Funding}

Not applicable. 

\subsection{Module Public Relations Funding}

Not applicable. 

\subsection{Module Standard Allowance for Gender Equality Measures}


Not applicable. 