% max 10 pages

 \subsection{Ethical and/or legal aspects of the project}
 
Ethical and legal considerations are not applicable to the proposed project since we 
will not conduct experiments on humans, animals, or genetic resources, nor will we  produce results that can be intentionally misused to cause harm. 

\todo{Update} We will conduct comprehension studies in Task 2.3 of the project to test the usefulness of our proposed solution. Such user studies are highly common in the context of computer science research. Participation in these studies will be fully voluntarily, participants will have the possibility to opt out at any moment, and we will not collect any personal data, as defined by the Ethics Committee of the University of Mannheim\footnote{\url{https://www.uni-mannheim.de/en/about/organization/bodies-and-committees/committees-and-councils/ethics-committee}}. Given that we will not collect any personal data, an ethics vote is not required according to the committee's guidelines.

%We would like to stress two ethical aspects with respect to the use of social media data. First, we will only use fully legal means (e.g, by using the official Twitter API) to obtain social media data. Second, we will replace the user names associated with the extracted posts with identifiers, such that the privacy of the users is protected (even though we will only use public posts). 

\subsection{Data handling}

\todo{Update} As explained in Section \ref{sec:objectives} and \ref{sec:wp6}, we plan to create new data sets consisting of event logs and related Twitter posts. Such data sets are not only relevant for this project, but also for other researchers in the area of process mining. In fact, the large interest in event logs has led to the creation of a central event log repository\footnote{https://www.tf-pm.org/resources/logs}, which is hosted and managed by the IEEE Task Force on Process Mining. As a member of the IEEE Task Force on Process Mining, we therefore plan to make our data sets publicly available via this repository. 

\subsection{Other information}

All relevant information is discussed above. 